Let's talk about Thurstone & Co.: An information-theoretical model for comparative judgments, and its statistical translation

### 2024-11-22 index.qmd
section:issues,CaseV; process:write

<!-- 6. What does the equal dispersion between stimuli imply? and does it hold? -->
In contrast, the assumption of *equal dispersion between stimuli* suggests that the variability in the perceived trait level of the stimuli is the same across stimuli. While Thurstone acknowledged that this assumption may be violated when "dealing with less conspicuous attributes or with less homogeneous stimuli" [@Thurstone_1927a, pp. 374], no study explicitly proposes that this assumption could also be violated due to the presence of an additional hierarchical (grouping) structure relevant to the texts. One such scenario might arise, for example, when comparing texts produced by university and secondary school students. In this case, university students may consistently (or more precisely) produce higher-quality texts, while secondary school students, who exhibit a broader range of writing abilities, would show greater variability in the quality of their texts. Although this example is somewhat contrived, it effectively illustrates how assuming equal dispersions across texts can overlook meaningful differences in the reliability of text quality across groups or individuals.

>>>>

<!-- 6. What does the equal dispersion between stimuli imply? and does it hold? -->
In contrast, the assumption of *equal dispersion between stimuli* suggests that the variability in the perceived trait level of the stimulus is the same across stimuli. While Thurstone acknowledged that this assumption may be violated when "dealing with less conspicuous attributes or with less homogeneous stimuli" [@Thurstone_1927a, pp. 374], no study explicitly proposes that this assumption could also be violated due to the presence of an additional hierarchical (grouping) structure relevant to the texts. One such scenario might arise, for example, when comparing texts produced by university and secondary school students. In this case, university students may consistently (or more precisely) produce higher-quality texts, while secondary school students, who exhibit a broader range of writing abilities, would show greater variability in the quality of their texts. Although this example is somewhat contrived, it effectively illustrates how assuming equal dispersions across texts can overlook meaningful differences in the reliability of text quality across groups or individuals.
